---
title: "Create your own score class object"
categories:
  - developer tools
type: learn-subsection
weight: 1
description: | 
 Create a new score class object for feature selection.
toc: true
toc-depth: 3
include-after-body: ../../../resources.html
---

```{r}
#| label: "setup"
#| include: false
#| message: false
#| warning: false
source(here::here("common.R"))
```

```{r}
#| label: "load"
#| include: false
#| message: false
#| warning: false
library(filtro)
library(modeldata)

pkgs <- c("filtro", "modeldata")
```

## Introduction

`r article_req_pkgs(pkgs)`

You can construct new scoring objects using `class_score()`. This article is a guide to creating new scoring objects. 

## Scoring object

`class_score` is the parent class of all subclasses related to the scoring method. There are a few properties for this object:

```{r}
#| label: "class_score"
library(filtro)
args(class_score)
```

-   `outcome_type`: What types of outcome can the method handle? The options are `numeric`, `factor`, or both. 

-   `predictor_type`: What types of predictor can the method handle? The options are `numeric`, `factor`, or both. 

-   `case_weights`: Does the method accpet case weights? `TRUE` or `FALSE`.

-   `range`: Are there known ranges for the statistic? For example, `c(0, Inf)`. 

-   `inclusive`: Are these ranges inclusive at the bounds? For example, `c(FALSE, FALSE)`. 

-   `fallback_value`: What is a value that can be used for the statistic so that it will never be eliminated? For example, `0`, `Inf`.

-   `score_type`: What is the column name that will be used for the statistic values? For example, `aov_pval`, `aov_fstat`. 

-   (Not used) `sorts`: How should the values be sorted (from most- to least-important)?

-   `direction`: What direction of values indicates the most important values? For example,  `maximum`, `minimize`.

-   `deterministic`: Does the fitting process use random numbers? `TRUE` or `FALSE`.

-   `tuning`: Does the method have tuning parameters? `TRUE` or `FALSE`.

-   `calculating_fn`: What function, if any, is used to estimate the values from data?

-   `label`: What label to use when printing? For example, `ANOVA p-values`, `ANOVA F-statistics`.

-   `packages`: What packages, if any, are required to train the method?

-   `results`: A slot for the results once the method is fitted. Initially, this is an empty data frame.

## Scoring object specific to the scoring method

As an example, let’s consider the ANOVA F-test filter. 

`class_score_aov` is a subclass of `class_score`. The subclass can also include additional properties specific to its implementation. For example: 

```{r}
# Create a subclass named 'class_score_aov'
class_score_aov <- S7::new_class(
  "class_score_aov",
  parent = class_score,
  properties = list(
    neg_log10 = S7::new_property(S7::class_logical, default = TRUE)
  )
)
```

In addition to the properties inherited from the parent class, `class_score_aov` also includes:

- `neg_log10`: Represent the score as `-log10(p_value)`? `TRUE` or `FALSE`.

For this filter, users can represent the score using either

- p-value or 

- F-statistic. 

We demonstrate how to create these instances (or objects) accordingly. 

`score_aov_pval` is an instance (or object) of the `class_score_aov` subclass, created using its constructor function: 

```{r}
# ANOVA p-value
score_aov_pval <-
  class_score_aov(
    outcome_type = c("numeric", "factor"),
    predictor_type = c("numeric", "factor"),
    case_weights = TRUE,
    range = c(0, Inf),
    inclusive = c(FALSE, FALSE),
    fallback_value = Inf,
    score_type = "aov_pval",
    direction = "maximize",
    deterministic = TRUE,
    tuning = FALSE,
    label = "ANOVA p-values"
  )
```

Individual properties can be accessed via `object@`. For examples: 

```{r}
score_aov_pval@case_weights
score_aov_pval@fallback_value
score_aov_pval@direction
```

`score_aov_fstat` is another instance (or object) of the `class_score_aov` subclass:  

```{r}
# ANOVA F-statistic
score_aov_fstat <-
  class_score_aov(
    outcome_type = c("numeric", "factor"),
    predictor_type = c("numeric", "factor"),
    case_weights = TRUE,
    range = c(0, Inf),
    inclusive = c(FALSE, FALSE),
    fallback_value = Inf,
    score_type = "aov_fstat",
    direction = "maximize",
    deterministic = TRUE,
    tuning = FALSE,
    label = "ANOVA F-statistics"
  )
```

## Fitting (or estimating) score

So far, we have discussed how to create a subclass and construct instances (objects) for the ANOVA F-test filter. 

The `fit()` function is a generic used to fit (or estimate) score.

```{r}
#| eval: false
score_aov_pval |>
  fit(Sale_Price ~ ., data = ames)
```

## Accessing results after fitting

Once the method is fitted via `fit()`, the data frame of results can be accessed via `object@results`. For examples: 

```{r}
library(modeldata)
ames_subset <- modeldata::ames |>
  # Use a subset of data for demonstration
  dplyr::select(
    Sale_Price,
    MS_SubClass,
    MS_Zoning,
    Lot_Frontage,
    Lot_Area,
    Street
  )
ames_subset <- ames_subset |>
  dplyr::mutate(Sale_Price = log10(Sale_Price))
```

```{r}
# # Specify ANOVA p-value and fit score
# ames_aov_pval_res <-
#   score_aov_pval |>
#   fit(Sale_Price ~ ., data = ames_subset)
# ames_aov_pval_res@results
```

```{r}
# # Specify ANOVA F-statistic and fit score
# ames_aov_fstat_res <-
#   score_aov_fstat |>
#   fit(Sale_Price ~ ., data = ames_subset)
# ames_aov_fstat_res@results
```

## Session information {#session-info}

```{r}
#| label: "si"
#| echo: false
small_session(pkgs)
```
