---
title: "Decision Tree Visualization"
categories:
  - model fitting
  - multivariate analysis
  - decision trees
type: learn-subsection
weight: 6
description: | 
  Visualize decision trees using different R packages and approaches.
  
toc: true
toc-depth: 2
r-packages:
  - tidymodels
  - rpart
  - partykit
  - rpart.plot
  - ggplot2
  - parttree
  - sessioninfo
include-after-body: ../../../html/resources.html
---

```{r}
#| label: "setup"
#| include: false
#| message: false
#| warning: false
source(here::here("R/common.R"))
```

```{r}
#| label: "load"
#| include: false
library(tidymodels)
library(rpart)
library(partykit)
library(rpart.plot)
library(parttree)
library(ggplot2)
library(sessioninfo)
pkgs <- unlist(rmarkdown::metadata[["r-packages"]])

theme_set(theme_bw() + theme(legend.position = "top"))
```


## Introduction

`r article_req_pkgs(pkgs)`

Decision trees are widely used because they are easy to interpret and can
capture nonlinear relationships between predictors and outcomes. Once a
tree model has been fit, visualization becomes an important step for
understanding splits, predictions, and decision boundaries.

Different R packages provide different approaches to tree visualization.
Some focus on simple structural displays, while others emphasize
publication-quality graphics or integration with the `ggplot2`
ecosystem.

This article compares and discusses several common methods for visualizing decision
trees in R:
  - `plot.rpart()`
  - `partykit::plot()`
  - `rpart.plot::rpart.plot()`
  - `parttree::geom_parttree()`

## Session information {#session-info}

```{r}
#| label: "si"
#| echo: false
small_session(pkgs)
```
 
